Bayesian determination of the effect of a deep euctectic solvent on the structure of lipid monolayers

### 2018-09-05 reports/paper.tex
cleaning everything up

\begin{table*}
	\small
	\caption{\ Lipid component volumes extracted from different literature sources. $V_l$ corresponds to the total lipid volume, MD to molecular dynamics simulation, WAXS to wide-angle X-ray scattering, NB to neutral buoyancy and DVTD to differential vibrating tube densimetry. $^a$ The values for the head component in Kucerka \emph{et al.},\cite{Kucerka2004} were taken from Balgav\'{y} \emph{et al}.\cite{Balgavy2001}}
	\label{tab:water}
	\begin{tabular*}{\textwidth}{@{\extracolsep{\fill}}l|lll|ll|ll|l|l}
		%\hline
    Lipid & DPPC & & & DMPC & & DLPC & & DMPG & POPG \\
    \hline
    Reference & [\citen{Armen1998}] & [\citen{Sun1994}] & [\citen{Kucerka2004,Balgavy2001}]$^a$ & [\citen{Armen1998}] & [\citen{Kucerka2004,Balgavy2001}]$^a$ & [\citen{Armen1998}] & [\citen{Kucerka2004,Balgavy2001}]$^a$ & [\citen{Pan2012}] & [\citen{Kucerka2012}] \\
    \hline
    $V_l$/\AA$^3$ & $1287.3\pm25.5$ & $1148\pm2$ & $1264.2\pm32.1$ & $1172.5\pm25.1£$ & $1155.4\pm30.0£$ & $1057.7\pm24.7$ & $1046.6\pm28.0$ & $1011.4$ & $1203$ \\
    $V_t$/\AA$^3$ & $966.4\pm5.4$ & $829\pm4$ & $924.7\pm17.6$ & $851.5\pm5.0£$ & $815.9\pm15.5£$ & $736.8\pm4.6$ & $707.1\pm13.5$ & $720.4$ & $914$ \\
    $V_h$/\AA$^3$ & $320.9\pm20.1$ & $319\pm6$ & $339.5\pm14.5$ & $320.9\pm20.1$ & $339.5\pm14.5$ & $320.9\pm20.1$ & $339.5\pm14.5$ & $291.0$ & $289$ \\
    Method & MD & WAXS & NB & MD & NB & MD & NB & DVTD & MD \\
    T/$^\circ$C & 50 & 24 & 30 & 50 & 30 & 50 & 30 & 20 & 25 \\
    %\hline
	\end{tabular*}
\end{table*}
In the XRR experiment, sample preparation was performed in situ using the standard method for the spreading of insoluble monolayers on water: a certain amount of the phospholipid solution was spread onto the liquid surface in order to provide a given surface concentration. After the evaporation of the chloroform, it is assumed that the resulting system is a solvent subphase with a monolayer of phospholipid at the interface. Surface concentration was modified by closing and opening the PTFE barriers of a Langmuir trough. In order to minimise the volumes used in the NR experiment (to keep the cost of deuterated compounds to a manageable level) it was not possible to use a Langmuir trough. Instead, small Delrin adsorption troughs were used that did not have controllable barriers. So, although the surface coverage was nominally the same as used in the X-ray studies, the lack of precise control over the surface pressure meant that it was not appropriate to co-refine XRR and NR contrasts together.

@@

In the first of two steps, this custom model was used to co-refine the component volume of the lipid head component, $V_h$, the volume of the tail component, $V_t$, and the head thickness, $d_h$ across XRR measurements at four different surface concentrations. The head thickness was not permitted to vary with surface pressure to reduce the dimensionality of the parameter space, an assumption that is based on our belief that increasing surface pressure should not significantly influence the head layer thickness. The following parameters were allowed to vary; $\theta_t$, and $\sigma_{t,h,s}$, independently across the surface pressures, while others, shown in Table \ref{tab:invariant}, were held constant at the values given. The length of the carbon chain was kept constant at the value determined by the Tanford equation,\cite{Tanford1980} since this is the maximum chain length possible and the most likely conformation at the surface pressures studied.\cite{Hollinshead2009} For each co-refinement of four XRR measurements, there were, in total, eleven degrees of freedom in the fitting process. Throughout all of the analysis, the reflectometry scale factor was allowed to vary freely, while the background as constrained to the intensity of either the largest or second-largest $q$-value.

@@

%
\begin{figure}
	\centering
	\includegraphics[width=0.48\textwidth]{figures/DLPC_all_data}
	\includegraphics[width=0.48\textwidth]{figures/DMPC_all_data}
	\includegraphics[width=0.48\textwidth]{figures/DPPC_all_data}
	\includegraphics[width=0.48\textwidth]{figures/DMPG_all_data}
	\caption{The XRR profiles (left) and SLD profiles (right) for each of the four lipids; (a) DLPC, (b) DMPC, (c) DPPC, (d) DMPG, at the four measured surface pressures; lowest surface pressure at the bottom and increasing moving up. The different surface pressure XRR profiles have been offset in the $y$-axis by an order of magnitude and SLD profiles offset in the $y$-axis by $5\times10^{-6}$ \AA$^{-2}$, for clarity. Source: Datasets, figure files and running/plotting scripts are available under CC-BY.\cite{mccluskey_2018}}
	\label{fig:lipids}
\end{figure}
%
\section{Results \& Discussion}
The chemically-consistent model was co-refined across the four surface pressure XRR measurements for each lipid. The resulting XRR profiles and associated SLD profiles are shown in Figure \ref{fig:lipids}. Table \ref{tab:liptab} gives details of all varied parameters for each lipid at the highest surface pressure measured, as well as the details of $d_t$ and $\phi_h$ which are determined from Eqns. \ref{equ:tl} and \ref{equ:phih} respectively. Each value is given with asymmetric uncertainies that corresponds to a 95 \% confidence interval of the PDF; the full PDF plots can be found in the ESI.

@@

%
\begin{figure}
	\centering
	\includegraphics[width=0.48\textwidth]{figures/DLPC_other_data}
	\includegraphics[width=0.48\textwidth]{figures/DMPC_other_data}
	\includegraphics[width=0.48\textwidth]{figures/DPPC_other_data}
	\includegraphics[width=0.48\textwidth]{figures/DMPG_other_data}
	\caption{The PDFs of the head component volume (left) and variation of $d_t$ (blue squares) and $\phi_h$ (green circles) with surface pressure for each of the four lipids; (a) DLPC, (b) DMPC, (c) DPPC, (d) DMPG. Source: Datasets, figure files and running/plotting scripts are available under CC-BY.\cite{mccluskey_2018}}
	\label{fig:lipresults}
\end{figure}
%
The variation of the tail layer thickness in the models with surface pressure is given for each lipid in Figure \ref{fig:lipresults}. It can be observed that as the surface pressure increases, the thickness of the tail layer increases to a point, before plateauing; for DPPC this occurs at 20 mNm$^{-1}$, DMPC at 30 mNm$^{-1}$ and for DMPG and DLPC can be assumed to be at higher pressures than those studied. This phenomenon of the tail thickness increasing with increasing surface pressure has been noted before for DMPC\cite{Bayerl1990} and DPPC\cite{Campbell2018} at the air-water interface.

@@

\subsection{Refinement of neutron reflectometry}
The ability to fit the NR data, as shown in Figure \ref{fig:neutron} indicates that the values found for the head component volume is consistant between the pair of measurements for the same system. It is clear, that again stable monolayers of the lipids are forming at the air-DES interface, and that the component volumes determined from XRR measurements are robust enough to be used in the modelling of NR data. Futhermore, the trends observed with increasing surface pressure in the XRR models, pertaining to the increasing tail thickness and decreasing solvent concentration in the head components are consistant with that found in the NR models.
%
\begin{figure}
	\centering
	\includegraphics[width=0.48\textwidth]{figures/nDMPC20_all_data}
	\includegraphics[width=0.48\textwidth]{figures/nDPPC20_all_data}
	\caption{The NR and SLD profiles at a surface pressure of 20 mNm$^{-1}$, the h-DES contrast is shown in blue while the hd-DES in green; (a) DMPC, (b) DPPC. The NR profiles have been offset in the $y$-axis by an order of magnitude and SLD profiles offset in the $y$-axis by $5\times10^{-6}$ \AA$^{-2}$, for clarity. Source: Datasets, figure files and running/plotting scripts are available under CC-BY.\cite{mccluskey_2018}}
	\label{fig:neutron}
\end{figure}
%
\begin{table}
	\small
	\caption{\ The best-fit values, and associated 95 \% confidence intervals for the varying parameters in the co-refined NR models. The values of $d_t$ were found from the appropriate values of $\theta_t$ using Eqn. \ref{equ:tl}, and the values of $\phi_h$ were found using Eqn. \ref{equ:phih}.}
	\label{tab:neutron}
	\begin{tabular*}{0.48\textwidth}{@{\extracolsep{\fill}}lllll}
		\hline
		Lipid & d$_{54}$-DMPC & & d$_{62}$-DPPC & \\
    SP/mNm$^{-1}$ & 20 & 25 & 15 & 20 \\
		\hline
		$\theta_t$/$^\circ$ & \input{../output/dmpc/angle20_neutron.txt} & \input{../output/dmpc/angle25_neutron.txt} & \input{../output/dppc/angle15_neutron.txt} & \input{../output/dppc/angle20_neutron.txt} \\
		$\sigma_{t,h,s}$/\AA & \input{../output/dmpc/rough20_neutron.txt} & \input{../output/dmpc/rough25_neutron.txt} & \input{../output/dppc/rough15_neutron.txt} & \input{../output/dppc/rough20_neutron.txt} \\
    \hline
    $\phi_h$/$\times10^{-2}$ & \input{../output/dmpc/solh20_neutron.txt} & \input{../output/dmpc/solh25_neutron.txt} & \input{../output/dppc/solh15_neutron.txt} & \input{../output/dppc/solh20_neutron.txt} \\
		$d_t$/\AA & \input{../output/dmpc/tail20_neutron.txt} & \input{../output/dmpc/tail25_neutron.txt} & \input{../output/dppc/tail15_neutron.txt} & \input{../output/dppc/tail20_neutron.txt} \\
		\hline
	\end{tabular*}
\end{table}
%

>>>>

\begin{table*}
	\small
	\caption{\ Lipid component volumes extracted from different literature sources. $V_l$ corresponds to the total lipid volume, MD to molecular dynamics simulation, WAXS to wide-angle X-ray scattering, NB to neutral buoyancy and DVTD to differential vibrating tube densimetry. $^a$ The values for the head component in Kucerka \emph{et al.},\cite{Kucerka2004} were taken from Balgav\'{y} \emph{et al}.\cite{Balgavy2001}}
	\label{tab:water}
	\begin{tabular*}{\textwidth}{@{\extracolsep{\fill}}l|lll|ll|ll|l|l}
		\hline
    Lipid & DPPC & & & DMPC & & DLPC & & DMPG & POPG \\
    \hline
    Reference & [\citen{Armen1998}] & [\citen{Sun1994}] & [\citen{Kucerka2004,Balgavy2001}]$^a$ & [\citen{Armen1998}] & [\citen{Kucerka2004,Balgavy2001}]$^a$ & [\citen{Armen1998}] & [\citen{Kucerka2004,Balgavy2001}]$^a$ & [\citen{Pan2012}] & [\citen{Kucerka2012}] \\
    \hline
    $V_l$/\AA$^3$ & $1287.3\pm25.5$ & $1148\pm2$ & $1264.2\pm32.1$ & $1172.5\pm25.1£$ & $1155.4\pm30.0£$ & $1057.7\pm24.7$ & $1046.6\pm28.0$ & $1011.4$ & $1203$ \\
    $V_t$/\AA$^3$ & $966.4\pm5.4$ & $829\pm4$ & $924.7\pm17.6$ & $851.5\pm5.0£$ & $815.9\pm15.5£$ & $736.8\pm4.6$ & $707.1\pm13.5$ & $720.4$ & $914$ \\
    $V_h$/\AA$^3$ & $320.9\pm20.1$ & $319\pm6$ & $339.5\pm14.5$ & $320.9\pm20.1$ & $339.5\pm14.5$ & $320.9\pm20.1$ & $339.5\pm14.5$ & $291.0$ & $289$ \\
    Method & MD & WAXS & NB & MD & NB & MD & NB & DVTD & MD \\
    T/$^\circ$C & 50 & 24 & 30 & 50 & 30 & 50 & 30 & 20 & 25 \\
    \hline
	\end{tabular*}
\end{table*}
In the XRR experiment, sample preparation was performed in situ using the standard method for the spreading of insoluble monolayers on water: a certain amount of the phospholipid solution was spread onto the liquid surface in order to provide a given surface concentration. After the evaporation of the chloroform, it is assumed that the resulting system is a solvent subphase with a monolayer of phospholipid at the interface. Surface concentration was modified by closing and opening the PTFE barriers of a Langmuir trough. In order to minimise the volumes used in the NR experiment (to keep the cost of deuterated compounds to a manageable level) it was not possible to use a Langmuir trough. Instead, small Delrin adsorption troughs were used that did not have controllable barriers. So, although the surface coverage was nominally the same as used in the X-ray studies, the lack of precise control over the surface pressure meant that it was not appropriate to co-refine XRR and NR contrasts together.

@@

In the first of two steps, this custom model was used to co-refine the component volume of the lipid head component, $V_h$, the volume of the tail component, $V_t$, and the head thickness, $d_h$ across XRR measurements at four different surface concentrations. In keeping with the work of Campbell \emph{et al.},\cite{Campbell2018} a single value for the head thickness was fitted for each lipid across all surface pressures, as the thickness of the head layer was considered to be dependent on molecular dimensions only. The following parameters were allowed to vary; $\theta_t$, and $\sigma_{t,h,s}$, independently across the surface pressures, while others, shown in Table \ref{tab:invariant}, were held constant at the values given. The length of the carbon chain was kept constant at the value determined by the Tanford equation,\cite{Tanford1980} since this is the maximum chain length possible and the most likely conformation at the surface pressures studied.\cite{Hollinshead2009} For each co-refinement of four XRR measurements, there were, in total, eleven degrees of freedom in the fitting process. Throughout all of the analysis, the reflectometry scale factor was allowed to vary freely, while the background as constrained to the intensity of either the largest or second-largest $q$-value.

@@

%
\begin{figure}
	\centering
	\includegraphics[width=0.45\textwidth]{figures/dlpc_ref_sld}
	\includegraphics[width=0.45\textwidth]{figures/dmpc_ref_sld}
	\includegraphics[width=0.45\textwidth]{figures/dppc_ref_sld}
	\includegraphics[width=0.45\textwidth]{figures/dmpg_ref_sld}
	\caption{The XRR profiles (left) and SLD profiles (right) for each of the four lipids; (a) DLPC, (b) DMPC, (c) DPPC, (d) DMPG, at the four measured surface pressures; lowest surface pressure at the bottom and increasing moving up. The different surface pressure XRR profiles have been offset in the $y$-axis by an order of magnitude and SLD profiles offset in the $y$-axis by $5\times10^{-6}$ \AA$^{-2}$, for clarity. Source: Datasets, figure files and running/plotting scripts are available under CC-BY.\cite{mccluskey_2018}}
	\label{fig:lipids}
\end{figure}
%
\section{Results \& Discussion}
The chemically-consistent model was co-refined across the four surface pressure XRR measurements for each lipid. The resulting XRR profiles and associated SLD profiles are shown in Figure \ref{fig:lipids}. Table \ref{tab:liptab} gives details of all varied parameters for each lipid at the highest surface pressure measured, as well as the details of $d_t$ and $\phi_h$ which are determined from Eqns. \ref{equ:tl} and \ref{equ:phih} respectively. Each value is given with asymmetric uncertainies that corresponds to a 95 \% confidence interval of the PDF; the full PDF plots can be found in the ESI.

@@

%
\begin{figure}
	\centering
	\includegraphics[width=0.45\textwidth]{figures/dlpc_vh_dt_phi}
	\includegraphics[width=0.45\textwidth]{figures/dmpc_vh_dt_phi}
	\includegraphics[width=0.45\textwidth]{figures/dppc_vh_dt_phi}
	\includegraphics[width=0.45\textwidth]{figures/dmpg_vh_dt_phi}
	\caption{The PDFs of the head component volume (left) and variation of $d_t$ (blue squares) and $\phi_h$ (green circles) with surface pressure for each of the four lipids; (a) DLPC, (b) DMPC, (c) DPPC, (d) DMPG. Source: Datasets, figure files and running/plotting scripts are available under CC-BY.\cite{mccluskey_2018}}
	\label{fig:lipresults}
\end{figure}
%
The variation of the tail layer thickness in the models with surface pressure is given for each lipid in Figure \ref{fig:lipresults}. It can be observed that as the surface pressure increases, the thickness of the tail layer increases to a point, before plateauing; for DPPC this occurs at 20 mNm$^{-1}$, DMPC at 30 mNm$^{-1}$ and for DMPG and DLPC can be assumed to be at higher pressures than those studied. This phenomenon of the tail thickness increasing with increasing surface pressure has been noted before for DMPC\cite{Bayerl1990} and DPPC\cite{Campbell2018} at the air-water interface.

@@

\subsection{Refinement of neutron reflectometry}
The ability to fit the NR data, as shown in Figure \ref{fig:neutron} indicates that the values found for the head component volume is consistant between the pair of measurements for the same system. It is clear, that again stable monolayers of the lipids are forming at the air-DES interface, and that the component volumes determined from XRR measurements are robust enough to be used in the modelling of NR data. Futhermore, the trends observed with increasing surface pressure in the XRR models, pertaining to the increasing tail thickness and decreasing solvent concentration in the head components are consistant with that found in the NR models.
%
\begin{figure}
	\centering
	\includegraphics[width=0.45\textwidth]{figures/nDMPC20_all_data}
	\includegraphics[width=0.45\textwidth]{figures/nDPPC20_all_data}
	\caption{The NR and SLD profiles at a surface pressure of 20 mNm$^{-1}$, the h-DES contrast is shown in blue while the hd-DES in green; (a) DMPC, (b) DPPC. The NR profiles have been offset in the $y$-axis by an order of magnitude and SLD profiles offset in the $y$-axis by $5\times10^{-6}$ \AA$^{-2}$, for clarity. Source: Datasets, figure files and running/plotting scripts are available under CC-BY.\cite{mccluskey_2018}}
	\label{fig:neutron}
\end{figure}
%
\begin{table}
	\small
	\caption{\ The best-fit values, and associated 95 \% confidence intervals for the varying parameters in the co-refined NR models. The values of $d_t$ were found from the appropriate values of $\theta_t$ using Eqn. \ref{equ:tl}, and the values of $\phi_h$ were found using Eqn. \ref{equ:phih}.}
	\label{tab:neutron}
	\begin{tabular*}{0.48\textwidth}{@{\extracolsep{\fill}}l|ll|ll}
		\hline
		Lipid & d$_{54}$-DMPC & & d$_{62}$-DPPC & \\
    SP/mNm$^{-1}$ & 20 & 25 & 15 & 20 \\
		\hline
		$\theta_t$/$^\circ$ & \input{../output/dmpc/angle20_neutron.txt} & \input{../output/dmpc/angle25_neutron.txt} & \input{../output/dppc/angle15_neutron.txt} & \input{../output/dppc/angle20_neutron.txt} \\
		$\sigma_{t,h,s}$/\AA & \input{../output/dmpc/rough20_neutron.txt} & \input{../output/dmpc/rough25_neutron.txt} & \input{../output/dppc/rough15_neutron.txt} & \input{../output/dppc/rough20_neutron.txt} \\
    \hline
    $\phi_h$/$\times10^{-2}$ & \input{../output/dmpc/solh20_neutron.txt} & \input{../output/dmpc/solh25_neutron.txt} & \input{../output/dppc/solh15_neutron.txt} & \input{../output/dppc/solh20_neutron.txt} \\
		$d_t$/\AA & \input{../output/dmpc/tail20_neutron.txt} & \input{../output/dmpc/tail25_neutron.txt} & \input{../output/dppc/tail15_neutron.txt} & \input{../output/dppc/tail20_neutron.txt} \\
		\hline
	\end{tabular*}
\end{table}
%
